Carbonate chemistry at Ny-Ålesund

### 2022-03-30 awipev-CO2.Rmd
refined stratification

sal_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=sal_delta)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous("Salinity", limits = c(-1.1, 0.8), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.65, label = substitute("S"["surface"] ~ "> S"["deep"])) +
  annotate("text", x = 1.5, y = -0.9, label = substitute("S"["surface"] ~ "< S"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(sal_delta_plt)

# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>% 
  group_by(depth_group2, montht) %>% 
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>% 
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>% 
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.3, label = substitute("T"["surface"] ~ "> T"["deep"])) +
  annotate("text", x = 1.5, y = -0.6, label = substitute("T"["surface"] ~ "< T"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>% 
  group_by(depth_group, montht)
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

@@

den_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=den_delta)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous("Density", limits = c(-1.5, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.1, label = substitute("D"["surface"] ~ "> D"["deep"])) +
  annotate("text", x = 1.5, y = -1.2, label = substitute("D"["surface"] ~ "< D"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(den_delta_plt)
# we look at pHint_tot_sf at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x = as.factor(depth_group), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf vs depth", x = "Depth (m)", y= "pH (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))

@@

pHinsi_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=pH_tot_sf_med_delta)) +
  #labs(y = substitute("pH"["T"] ~ "at in situ temp."), x = "Month") +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(substitute("pH"["T"] ~ "at in situ temp."), limits = c(-0.07, 0.03), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.02, label = substitute("pH"["surface"] ~ "> pH"["deep"])) +
  annotate("text", x = 1.5, y = -0.06, label = substitute("pH"["surface"] ~ "< pH"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(pHinsi_delta_plt)
# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

>>>>

sal_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=sal_delta)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Salinity", limits = c(-1.1, 0.8), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.65, label = substitute("S"["surface"] ~ "> S"["deep"])) +
  annotate("text", x = 2, y = -0.9, label = substitute("S"["surface"] ~ "< S"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(sal_delta_plt)

# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>% 
  group_by(depth_group2, montht) %>% 
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>% 
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>% 
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, label = substitute("T"["surface"] ~ "> T"["deep"])) +
  annotate("text", x = 2, y = -1.8, label = substitute("T"["surface"] ~ "< T"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>% 
  group_by(depth_group, montht)
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

@@

den_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=den_delta)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Density", limits = c(-1.5, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.25, label = substitute("D"["surface"] ~ "> D"["deep"])) +
  annotate("text", x = 2, y = -1.2, label = substitute("D"["surface"] ~ "< D"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(den_delta_plt)
# we look at pHint_tot_sf at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x = as.factor(depth_group), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf vs depth", x = "Depth (m)", y= "pH (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))

@@

pHinsi_delta_plt <- tmp_delta %>% 
  ggplot(aes(x=montht, y=pH_tot_sf_med_delta)) +
  #labs(y = substitute("pH"["T"] ~ "at in situ temp."), x = "Month") +
  scale_x_discrete("Month") +
  scale_y_continuous(substitute("pH"["T"] ~ "at in situ temp."), limits = c(-0.07, 0.03), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 1.5, y = 0.02, label = substitute("pH"["surface"] ~ "> pH"["deep"])) +
  annotate("text", x = 1.5, y = -0.06, label = substitute("pH"["surface"] ~ "< pH"["deep"])) +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(pHinsi_delta_plt)
# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))
